---
title: "Introduction to handy"
author: "Kevin Pan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to handy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Philosophy 

There are common steps that you may face in your data exploration:

- Import (messy) data
- **Clean** messy data
- **Describe** cleaned data
- **Plot** cleaned data
- Model data
- **Write** output tables

This package aims to facilitate this basic data handling with functions that I've found handy. 

## Quick Reference Table

| handy Function    | Use Case     | Description |
|-------------------|--------------|----------------------|
| `clean_names`     |  Clean Data      | Normalize variable names of a data.frame to snake_case |
| `title_names`     |  Clean Data      | Normalize variable names of a data.frame to Title Case |
| `remove_columns`  |  Clean Data      | Removes columns that are either missing or non-unique |   
| `add_mean_row`    |  Describe Data   | Add a mean row to a data.frame of numeric values |
| `add_total_row`   |  Describe Data   | Add a total row to a data.frame of numeric values |
| `check_variables` |  Describe Data   | Calculate % missing and number of unique values for all variables |
| `%p%`             |  Shorthand       | Allow string concatenation by piping |
| `cs`              |  Shorthand       | Makes a character vector without quotes |
| `named_list`      |  Shorthand       | Makes a named list using object names |
| `multiplot`       |  Plotting        | Arrange multiple plots |
| `scale_x_human`   |  Plotting        | Scales ggplot axis to easily readable K, M, B, etc. |
| `scale_y_human`   |  Plotting        | Scales ggplot axis to easily readable K, M, B, etc. |
| `write_excel`     |  Write Output    | Write a list of data.frames in R to named sheets in an Excel workbook |
| `write_regression_to_excel` | Write Output | Write a list of data.frames in R to named sheets in an Excel workbook |


## Clean 

Let's use R's built in `iris` dataset. 

These cleaning functions are built to be compatible with magrittr/dplyr pipes. 

```{r}
library(handy)
library(magrittr)
library(knitr)
iris3obs <- head(iris, 3) 
iris3obs %>% kable()
iris3obs %>% clean_names %>% kable()
```

What if we want to focus on columns that change for the set we're looking for?

```{r}
iris3obs %>% remove_columns(1) %>% kable()
```

## Describe

For every dataset, we want to know what it contains: 
```{r}
mtcars %>% check_variables %>% kable
```

Sometimes it's helpful to add a mean or total row for a table that's to be printed out. 
These are best for datasets with a character key as the first column. 
```{r}
arrests <- USArrests %>% tibble::rownames_to_column() %>% dplyr::as_tibble()
arrests %>% head %>% add_mean_row %>% kable
```

```{r}
arrests %>% head %>% add_total_row %>% kable
```

## Plot 

```{r}
library(ggplot2)
p1 <- Seatbelts %>% tibble::as_tibble() %>% 
  ggplot(aes(kms, DriversKilled)) +
  geom_point() +
  scale_x_human() 
p2 <- Seatbelts %>% tibble::as_tibble() %>% 
  ggplot(aes(DriversKilled)) +
  geom_histogram(alpha = .5, bins = 30)
multiplot(p1, p2)
```

## Write

And for some analyses, it's nice to write out the final data to Excel.

```{r, eval=FALSE}
named_list(mtcars, iris) %>% 
  write_excel(workbook_fname = 'wb.xlsx')

readxl::excel_sheets('wb.xlsx')
readxl::read_excel('wb.xlsx', sheet = 'iris') %>% head %>% kable
```

Here, using `named_list` assembles a list with the list names set to the names of the objects.

```{r}
lm_assault <- lm(Assault ~ UrbanPop, data = arrests)
lm_anscombe <- lm(y1 ~ x1, data = anscombe)
reg_list <- named_list(lm_assault, lm_anscombe)

reg_list
```

```{r, eval=FALSE}
reg_list %>% write_regression_to_excel('reg.xlsx')
readxl::read_excel('reg.xlsx', sheet = 'lm_anscombe') %>% head %>% kable
```

## Other

What do we have in memory at this point?
```{r}
lsos(n = 5) %>% kable
```

